### 2022-11-26 dissertation/chapters/06-bayesian-hyper-heuristic.tex
🚧 (final-review): Chapter 5 WIP

\chapter{Bayesian Hyper-Heuristic}\label{chap:bhh}

\begin{quotation}
      ``The result is a posterior distribution, which the agent may use as its new prior in the next step.''
\end{quotation}
\begin{flushright}
      - Pedro Domingos, The Master Algorithm
\end{flushright}

The above quote was the inspiration for the development of a novel \acf{HH} that uses Bayesian probability concepts as a selection mechanism to drive the heuristic selection process. Thus far the reader has been presented with all of the necessary background information on \acp{ANN} (Chapter \ref{chap:anns}), low-level \index{heuristic}heuristics (Chapter \ref{chap:heuristics}), \acp{HH} (Chapter \ref{chap:hhs}) and lastly, probability theory (Chapter \ref{chap:probability}). These elements form the fundamental components of the proposed \Acf{BHH}. This chapter provides the details around the implementation of the \Ac{BHH} and explains how it used to train \acp{FFNN}. The remainder of the chapter is structured as follows:

\begin{itemize}
      \item \textbf{Section \ref{sec:bhh:overview}} provides a brief overview of the \Acs{BHH}.

      \item \textbf{Section \ref{sec:bhh:architecture}} provides the general architecture and \Ac{HH} framework implemented by the \Acs{BHH}.

      \item \textbf{Section \ref{sec:bhh:heuristic_pool}} presents the \index{heuristic}\textit{heuristic pool}, a collection of low-level \index{heuristic}heuristics. Discussions follow on the importance of diversity amoungst heuristics in the \index{heuristic pool}heuristic pool.

      \item \textbf{Section \ref{sec:bhh:entity_pool}} presents details on entity (local) and population (global) memory (state).

      \item \textbf{Section \ref{sec:bhh:performance_log}} presents detailed discussions on performance measurement. The \textit{performance log}, implemented by the \acs{BHH}, is presented.

      \item \textbf{Section \ref{sec:bhh:credit_assignment_strategy}} presents detailed discussions on credit assignment and move acceptance strategies.

      \item \textbf{Section \ref{sec:bhh:selection_mechanism}} presents the Bayesian probabilistic model that is used as the heuristic selection mechanism.

      \item \textbf{Section \ref{sec:bhh:optimisation_step}} presents the learning mechanisms by which the probabilistic model can be optimised.

      \item \textbf{Section \ref{sec:bhh:hyper_parameters}} summarises and discusses the associated \index{hyper-parameters}hyper-parameters and default values.

      \item \textbf{Section \ref{sec:bhh:algorithm}} provides the pseudo-code algorithm for the \acs{BHH}.

      \item \textbf{Section \ref{sec:bhh:summary}} provides a brief summary of the chapter.
\end{itemize}

\section{Overview}\label{sec:bhh:overview}

This section provides an overview of the workings of the \Acf{BHH}. The general concept of the \acs{BHH} can be summarised as follows. The \acs{BHH} implements a high-level heuristic selection mechanism that learns to select the best heuristic, from a pool of low-level heuristics, to apply to a population of entities, each implementing a candidate solution to a \acs{FFNN}, with the intent of both optimising the underlying \acs{FFNN} and \acs{FFNN} training process. The \acs{BHH} does so by learning the probability that a given heuristic will perform well at a given stage in the \acs{FFNN} training process. These probabilities are then used as heuristic selection probabilities in the next step of the training process.

Formal classification of the \Acs{BHH} is needed. Chapter \ref{chap:hhs} presented the reader with a proposed classification scheme for \acp{HH} by~\citeauthor{ref:burke:2010}~\cite{ref:burke:2010}. According to the aforementioned classification scheme, the \Acs{BHH} is a population-based, meta-hyper-heuristic that utilises selection and perturbation of low-level heuristics in an online learning fashion. A breakdown of the classification is given as follows:

\begin{itemize}
      \item \textbf{Population-Based:} The \acs{BHH} implements a population-based approach, where a collection of different candidate solutions, referred to as \textit{entities}, work together to yield a global best solution.

      \item \textbf{Meta-Hyper-Heuristic:} There exists a domain barrier, where the \acs{BHH} searches through the \textit{heuristic space}, using only \index{heuristic}heuristic performance information, while lower level heuristics search through the \textit{solution space} using information from the search domain itself. Furthermore, the \index{heuristic pool}heuristic pool implemented by the \acs{BHH} supports both gradient-based low level \index{heuristic}heuristics, as well as meta-heuristics.

      \item \textbf{Selection and Perturbation of Low-Level Heuristics:} The \acs{BHH} implements a \index{heuristic}heuristic selection mechanism that selects from a collection of lower level \index{heuristic}heuristics, called a \index{heuristic pool}\textit{heuristic pool}. Selection is biased towards good performing heuristics. A credit assignment strategy is used to \textit{reward} good heuristic performance. The \acs{BHH} maintains entity and population state through operations that \textit{proxy} update steps from different heuristics (perturbation), as is required. These proxy operations ensure that heuristic requirements are satisfied even when different heuristics are used for different entities throughout the training process.

      \item \textbf{Online Learning:} The \acs{BHH} applies learning, at specified intervals, throughout the \acs{FFNN} training process.
\end{itemize}

The following sections presents the reader with more detail on the implementation of the \acs{BHH}. Discussions lead with the \acs{BHH} architecture and general \acs{HH} framework in the following section.

\section{Architecture}\label{sec:bhh:architecture}

This section aims to present the reader with all the high level components in the architecture of the \Acs{BHH}. Burke et al.~\cite{ref:burke:2010} proposed an initial framework for \acp{HH} and Grobler~\cite{ref:grobler:2015} further proposed a framework for a heterogeneous meta-\ac{HH}. The aforementioned frameworks are adapted for the implementation of the \Acs{BHH}. An illustration of the high-level architecture of the \Acs{BHH} is given in Figure \ref{fig:bhh_architecture}.

\begin{figure}[htbp]
      \centering
      \includegraphics[width=1.0\textwidth]{bhh_architecture.pdf}
      \caption[An illustration of
            the architecture and high level components of the \acf{BHH}.]{An illustration of
            the architecture and high level components of the \acf{BHH}.}
      \label{fig:bhh_architecture}
\end{figure}

With reference to Figure \ref{fig:bhh_architecture}, the components are briefly given below in the order of information flow during the \acs{FFNN} training process steps.

\begin{itemize}
      \item \textbf{Initialisation Step}: The initialisation step refers to the initialisation strategy implemented by the \acs{BHH}.

      \item \textbf{Heuristic Pool}: The \index{heuristic pool}heuristic pool contains a collection of low-level \index{heuristic}heuristics.

      \item \textbf{Entity Pool}: The \index{entity pool}entity pool represents a collection of candidate solutions for the underlying \acs{FFNN}, referred to as entities in a population.

      \item \textbf{Selection Mechanism}: The selection mechanism refers to the Bayesian probabilistic model.

      \item \textbf{Heuristic-Entity Selections}: Every entity in the \index{entity pool}entity pool gets assigned a selected heuristic.

      \item \textbf{Train and Test Datasets}: The training set is used to train the model while the test set is used to evaluate the model's generalisation capabilities.

      \item \textbf{Model}: The underlying \acs{FFNN} to be trained.

      \item \textbf{Low-Level Heuristics Update Step}: The low-level heuristics' update step as it applies to its allocated entity.

      \item \textbf{Loss Function}: In supervised learning, the loss function is a measure of distance between predicted output and the ground truth, and is the mechanism used to evaluate the model's performance.

      \item \textbf{Domain Barrier}: The domain barrier is the logical separation of information available in the heuristic space and the solution space. The \acs{BHH} searches in the heuristic space, while low-level heuristics search in the solution space.

      \item \textbf{Performance Log}: Contains a record of heuristic-entity performance at each step of the training process.

      \item \textbf{Credit Assignment Strategy}: The strategy used to assign credit to heuristics for their performance.

      \item \textbf{Update Step}: The high-level heuristic update step as implemented by the \acs{BHH}.
\end{itemize}

Each of the components in Figure \ref{fig:bhh_architecture} is discussed in detail in the following sections.

\index{heuristic pool}
\section{Heuristic Pool}\label{sec:bhh:heuristic_pool}

Generally speaking, the \index{heuristic pool}heuristic pool is a collection of low-level heuristics under consideration by the \acs{BHH}. The \index{heuristic pool}heuristic pool contains the set of low-level \index{heuristic}heuristics that, together with their performance information, make up the \text{heuristic space}. Importantly, the \acs{BHH} searches in \index{heuristic}heuristic space. The \index{heuristic pool}heuristic pool is defined in terms of the diversity of heuristics, as well as the amount of heuristics in the \index{heuristic pool}heuristic pool. Each of these are discussed in the following sections.

\subsection{Heuristic Diversity}\label{sec:bhh:heuristic_pool:diversity}

A \index{heuristic}heuristic's eligibility for inclusion in the \index{heuristic}heuristic pool, is determined by its ability to solve to underlying problem and its search behaviour and characteristics. The \index{heuristic pool}heuristic pool should contain a variety of different heuristics that have different search behaviours, so that a trade-off can be made by between exploration and exploitation of solutions, as is required during the training process. Heuristics that explore a lot, should be selected when exploration is needed, and heuristics that exploit a lot should be selected accordingly. The trade-off between exploration and exploitation is managed by the high-level heuristic. The \acs{BHH} will learn to select and apply the appropriate heuristic, at the appropriate points, throughout the training process. In doing so, a balance is achieved between exploration and exploitation.

\subsection{Heuristic Pool Size}\label{sec:bhh:heuristic_pool:pool_size}

Another design decision in the heuristic pool is heuristic pool size. Since the \acs{BHH} is a \textit{learning} \acs{HH}, every heuristic that is included in the heuristic pool increases the requirement of more statistical evidence of heuristic performance. Not only does a large \index{heuristic pool}heuristic pool drastically complicate the learning process that is required by the \acs{BHH}, but it also drastically complicates the process of maintaining \textit{state}. State is the term used to describe data that is relevant to the training process, and is maintained between training steps. An example of \textit{local} state is an entity's position, while an example of \textit{global} state is the global best solution found thus far. The following section presents the concept of \textit{proxies} and their role in maintaining state.

\subsection{Proxies}\label{sec:bhh:heuristic:proxies}

The concept of proxies arise from the sparsity of state as maintained by different \index{heuristic}heuristics. Since \index{heuristic}heuristics maintain (possibly) different states, there is an uncertainty of state transition when switching between \index{heuristic}heuristics. Consider an example where the \index{heuristic pool}heuristic pool consists of just two \index{heuristic}heuristics. One heuristic is a gradient-based heuristic that maintains momentum such as \acs{Adam}. The other is a \index{meta-heuristic}meta-heuristic that does not require a gradient, such as \acs{PSO}. Both these \index{heuristic}heuristics track different parameters in their state. For \acs{Adam}, the expected gradient mean and variance is maintained, while the \ac{PSO} maintains record of the gbest and pbest solutions of all of entities. A solution to state indifference, is to proxy heuristic state update operations. State is then maintained in two parts:

\begin{itemize}
      \item \textbf{Primary State}: Refers to the state that is originally maintained by a heuristic. The selected heuristic simply applies the normal state update operations to its state.

      \item \textbf{Proxied State}: Refers to the state that is not directly maintained by the heuristic, but can be updated by outsourcing the required state update operation to another heuristic.
\end{itemize}

Primary and proxied state parameters must be maintained together. Since entities represent candidate solutions, which are forms of state, entities are ideal candidates to store and maintain these state parameters. Entities are extended to include the primary state elements of \textit{all} the underlying low-level heuristics, as well as the candidate solution (weights) for the \acs{FFNN}. At each heuristic-entity application step, all state parameters, per entity, are updated either by primary method or by proxied method. The \acs{BHH} thus incorporates a mapping of proxied state update operation as given in the example in Table \ref{tab:bhh:heuristic_pool:proxy_mapping_example} below.

\begin{table}[htbp]
      \centering
      \caption{An example of mapping of proxied state update operation maintained by the \acs{BHH}.}
      \label{tab:bhh:heuristic_pool:proxy_mapping_example}%
      \par\bigskip
      \resizebox{0.5\textwidth}{!}{
            \begin{tabular}{ccccc}
                                                                  &   & \multicolumn{3}{c}{State Parameter}                                                                                                                                                                                      \\
                  \cmidrule{3-5}                                  &   & 1                                                                      & 2                                                                      & 3                                                                      \\
                  \cmidrule{3-5}    \multirow{3}[1]{*}{Heuristic} & A & \cellcolor[rgb]{ .776,  .937,  .808}\textcolor[rgb]{ 0,  .38,  0}{n/a} & \cellcolor[rgb]{ 1,  .922,  .612}\textcolor[rgb]{ .612,  .341,  0}{B}  & \cellcolor[rgb]{ .776,  .937,  .808}\textcolor[rgb]{ 0,  .38,  0}{n/a} \\
                                                                  & B & \cellcolor[rgb]{ .776,  .937,  .808}\textcolor[rgb]{ 0,  .38,  0}{n/a} & \cellcolor[rgb]{ .776,  .937,  .808}\textcolor[rgb]{ 0,  .38,  0}{n/a} & \cellcolor[rgb]{ 1,  .922,  .612}\textcolor[rgb]{ .612,  .341,  0}{A}  \\
                                                                  & C & \cellcolor[rgb]{ .776,  .937,  .808}\textcolor[rgb]{ 0,  .38,  0}{n/a} & \cellcolor[rgb]{ 1,  .922,  .612}\textcolor[rgb]{ .612,  .341,  0}{B}  & \cellcolor[rgb]{ 1,  .922,  .612}\textcolor[rgb]{ .612,  .341,  0}{A}  \\
            \end{tabular}%
      }
\end{table}%

From the example given in Table \ref{tab:bhh:heuristic_pool:proxy_mapping_example}, when \index{heuristic}heuristic A is selected, it will outsource state update operations from \index{heuristic}heuristic B, for state parameter 2. Heuristic B will outsource from \index{heuristic}heuristic A, for state parameter 3. Finally, \index{heuristic}heuristic C will outsource from \index{heuristic}heuristic A and B, for state parameters 2 and 3 respectively. In this way, all \index{heuristic}heuristics maintain all the state parameters.

Proxied state update operations is a simple concept in principle, but requires detailed decomposition of the \index{heuristic}heuristics included in the \index{heuristic pool}heuristic pool.  Overlapping and unique state parameters must be identified so that a proxy mapping, such as the one given in Table \ref{tab:bhh:heuristic_pool:proxy_mapping_example}, can be constructed. A suggestion to simplify this process is to borrow concepts from the equations of motion from physics. These include \textit{position}, \textit{velocity}, \textit{acceleration} and \textit{momentum}. Expressing \index{heuristic}heuristic update steps according to these parameters drastically simplify the process. However, it is possible that \index{heuristic}heuristics implement unique state parameters that do not overlap, these have to be catered for in the proxy mapping.

State parameters and update operations should not be considered in isolation. An example is \textit{position} and \textit{velocity}. Consider, for example, \index{heuristic}heuristics such as \acf{DE} and \acfp{GA}. These \index{heuristic}heuristics recombine entities. In the aforementioned case, the concept of an equation of motion does not entirely make sense, since the displacement of its position is not a result of maintaining velocity or momentum, but rather by displacement through recombination. In this particular case, a solution is to apply the recombination operation to \textit{all} applicable state parameters as well, or to nullify the necessary state parameters. Unfortunately there is no general, automated solution to formulate these heuristic state overlaps. Each \index{heuristic}heuristic must be carefully considered. More details on the \index{entity pool}entity pool and entity state are presented in the following section.

\index{entity pool}
\section{Entity Pool}\label{sec:bhh:entity_pool}

The \index{entity pool}entity pool refers to a collection of \textit{entities} that each represent a candidate solution to the underlying \acs{FFNN} being trained. The entities contain information of the solution space. A common naming convention for such a collection is a \textit{population} of entities. The \acs{BHH} is a population-based \acs{HH} and as such, the \index{entity pool}entity pool size or population size is an important design choice. The population size is defined as a hyper-parameter of the \acs{BHH} and can be empirically evaluated.

The \index{entity pool}entity pool maintains two different types of state. These include entity (local) and population (global) state. Each of these are discussed in more detail next.

\subsection{Entity State}\label{sec:bhh:entity_pool:entity_state}

Entities represent candidate solutions to the model's trainable parameters (weights) and other \index{heuristic}heuristic-specific state parameters, as was discussed in Section \ref{sec:bhh:heuristic:proxies}. It can be said that entities implement \textit{local} state. It was mentioned that these entities can be treated as physical \textit{particles} in a hyper-dimensional search environment. Entities model concepts from physics. For example, the candidate solution is represented as the entity's position, and an entity's velocity and acceleration is analogous to the gradient and momentum of the entity respectfully. Examples of entity state parameters, as derived from various low-level \index{heuristic}heuristics, is given as follows.

\begin{itemize}
      \item \textbf{position}: General parameter that represents the actual candidate solution and is thus a primary state parameter for all \index{heuristic}heuristics.

      \item \textbf{velocity}: Directly implemented by \index{heuristic}heuristics such as \acp{PSO} and is analogous to the gradient for gradient-based \index{heuristic}heuristics such as \ac{Momentum}.

      \item \textbf{gradient}: The last known gradient, as derived from gradient-based \index{heuristic}heuristics.

      \item \textbf{position delta}: The last computed position delta between the current timestep and the previous timestep.

      \item \textbf{sum of the gradients squared}: As required and maintained by \index{heuristic}heuristics such as \acs{Adagrad}.

      \item \textbf{expected position delta variance}: As required and maintained by \index{heuristic}heuristics such as \acs{Adadelta}.

      \item \textbf{expected gradient mean}: As required and maintained by \index{heuristic}heuristics such as \acs{Momentum}, \acs{NAG} and \acs{Adam}.

      \item \textbf{expected gradient variance}: As required and maintained by \index{heuristic}heuristics such as \acs{RMSProp}, \acs{Adadelta} and \acs{Adam}.

      \item \textbf{personal best position}: Parameter that tracks that best known position by the entity thus far. As required and maintained by \index{heuristic}heuristics such as \acs{PSO}.

      \item \textbf{personal best loss}: Parameter that tracks that best known loss by the entity thus far. As required and maintained by \index{heuristic}heuristics such as \acs{PSO}.

>>>>

\chapter{Probability}
\label{chap:probability}

\begin{quotation}
      ``Probability theory is nothing but common sense reduced to calculation.''
\end{quotation}
\begin{flushright}
      - Pierre-Simon Laplace
\end{flushright}

Probability theory and statistics can be traced back to as early as the 18th century. In 1718,~\citeauthor{ref:demoivre:1718}~\cite{ref:demoivre:1718} published the \textit{Doctrine of of Chance}; a book that is widely seen as the first published book on probability theory. In 1763, Thomas Bayes~\cite{ref:bayes:1763} published an article titled \textit{An Essay towards solving a Problem in the Doctrine of Chances} where the first version of \index{Bayes' theorem}Bayes' theorem was introduced.

Probability theory, statistics and \acf{ML} are transdisciplinary fields with many shared concepts. There are many examples of how probability theory has been incorporated into \acs{ML} research. In 1991,~\citeauthor{ref:denker:1991},~\cite{ref:denker:1991} proposed a way to transform \acf{ANN} outputs to probability distributions. In 1993, \citeauthor{ref:neal:1993}~\cite{ref:neal:1993} developed a \acf{MCMC} sampling algorithm for Bayesian \acp{NN}. These are but a few examples of the role that probability theory has played in \acs{ML} research in the past.

Chapter \ref{chap:hhs} provided the reader with the concept of a \acf{HH}. This dissertation aims to develop a selection \acs{HH} that makes use of probability theory to select the best \acs{HH} to solve a given problem. This chapter aims to provide the reader with the necessary background information on probability theory and statistics. These are big fields and focus is put on the elements that are required to formulate the proposed \acf{BHH}. The remainder of the chapter is structured as follows:

\begin{itemize}
      \item \textbf{Section \ref{sec:probability:overview}} gives a brief overview of what probability is and how it is used.

      \item \textbf{Section \ref{sec:probability:multiple_events}} presents the concept of conditional probability.

      \item \textbf{Section \ref{sec:probability:multiple_events}} presents the two laws of probability related to the intersection and union of multiple events.

      \item \textbf{Section \ref{sec:probability:cond_probability}} introduces conditional probability.

      \item \textbf{Section \ref{sec:probability:bayes_theorem}} introduces \index{Bayes' theorem}Bayes' theorem, the fundamental theorem upon which the \acs{BHH} is built.

      \item \textbf{Section \ref{sec:probability:probability_distributions}} presents the reader with relevant probability distributions.

      \item \textbf{Section \ref{sec:probability:conjugate_priors}} presents the reader with relevant conjugate prior probability distributions.

      \item \textbf{Section \ref{sec:probability:bayesian_statistics}} presents \textit{Bayesian} statistics. Brief discussions follow on the \textit{frequentist} view and \textit{Bayesian} view of probability. Detailed discussions follow on Bayesian optimisation methods such as \index{Bayesian analysis}\textit{Bayesian analysis}.

      \item \textbf{Section \ref{sec:probability:summary}} provides a brief summary of the chapter.
\end{itemize}

\section{Overview of Probability}\label{sec:probability:overview}

In everyday conversation, the term \textit{probability} is a measure of one's belief in the occurrence of a future event~\cite{ref:wackerly:2014}. Probability is a necessary tool used in many fields including physics, biology, chemistry and computer science. These fields contain many cases that generate observations that cannot be predicted with absolute certainty~\cite{ref:wackerly:2014}. Probability can be be inferred and confirmed through past events. These events are referred to as \textit{random} or \textit{stochastic} events. The probability that a certain event $A$ might occur, is denoted $P(A)$. Although these random events cannot be predicted with absolute certainty, the relative frequency with which they occur over many trials, is often remarkably stable.

Consider flipping an unbiased, fair coin. The coin has two possible outcomes. It can conclude that each side has a $\frac{1}{2}$ or $50\%$ chance of occurring. In statistics, the decimal probability notation is used, where $0 <= P(A) <= 1$. Suppose the fair coin is thrown 10 times, one is not guaranteed to observe $0.5$ heads and $0.5$ tails. There is some probability (although small) that the coin might fall on heads $0/10$ times. The probability of such an event occurring is $0.0009765625$. In the coin flip example, the \acf{CLT} shows that the normalised sum of events tends toward a normal distribution with a mean value of $0.5$ if the number of events observed $N$, is large~\cite{ref:wackerly:2014}. The larger the value of $N$, the higher the confidence of mean probability and relative frequency of the event. The stable long-term relative frequency by which a random event occurs, provides an intuitive and meaningful measure of belief that a certain event will occur again at some point in the future~\cite{ref:wackerly:2014}.

Probability can also be expressed over multiple random events. Multiple random events can be considered together, dependently or conditionally. The following sections provides insight into conditional and joints probabilities of multiple random events.

\section{Conditional Probability and Independence}\label{sec:probability:cond_probability}

The occurrence of a given random event $A$ can often be conditional on the occurrence of another event $B$. In the field of medicine, an example of this is to calculate the probability of a certain diagnosis of a sick patient given his/her symptoms. The aforementioned relationship between events is referred to as the conditional probability between two events. The conditional probability is expressed as $P(A \vert B)$ and is read \textit{the probability of $A$ given $B$}. On the contrary, the \textit{unconditional} probability is the probability of an event, not dependent on any other. The conditional probability of $A$ given $B$ can be expressed as is given in Definition \ref{eq:probability:cond_probability} below~\cite{ref:wackerly:2014}.
\\
\begin{definition}[\textbf{Conditional Probability}]
      \label{eq:probability:cond_probability}
      The conditional probability of an event $A$, subject to the occurrence of event $B$ is expressed as

      \begin{equation}
            \label{eq:probability:overview:conditional}
            P(A \vert B) = \frac{P(A \cap B)}{P(B)}
      \end{equation}
\end{definition}

In Equation~\eqref{eq:probability:overview:conditional} above, $P(B) > 0$. Note that conditional probability does not suggest causation. If an event $A$ has a high probability of occurring after observing event $B$, it does not necessarily mean that $A$ is caused by $B$. Conditional probability simply expresses the dependence amongst events.

It could also be the case that the outcome of observing event $A$ is not affected by the occurrence of $B$. In this case, it is said that events $A$ and $B$ are independent. The independence between two events are expressed in Definition \ref{def:probability:cond_probability:independence} below.
\\
\begin{definition}[\textbf{Independence of Events}]
      \label{def:probability:cond_probability:independence}
      Two events, $A$ and $B$ are said to be independent of each other if, and only if the following criteria hold:

      \begin{itemize}
            \item $P(A \vert B) = P(A)$
            \item $P(B \vert A) = P(B)$
            \item $P(A \cap B) = P(A)P(B)$
      \end{itemize}

      otherwise, events $A$ and $B$ are said to be \textit{dependent} random events.
\end{definition}

\section{Two Laws of Probability for Multiple Events}\label{sec:probability:multiple_events}

Suppose there are two random events $A$ and $B$, then one can calculate the probability of the union and intersection of these events. From the aforementioned concept, two laws of probability can be formulated. These are referred to as the \textit{multiplicative} and \textit{additive} laws of probability~\cite{ref:wackerly:2014} and is given below in Theorems \ref{th:probability:multiple_events:multiplicative} and \ref{th:probability:multiple_events:additive} respectively.\\

\begin{theorem}[\textbf{The Multiplicative Law of Probability}]
      \label{th:probability:multiple_events:multiplicative}
      The probability of the intersection of two events $A$ and $B$ is given as

      \begin{equation}
            \begin{split}
                  P(A \cap B)
                  &= P(A)P(B \vert A) \\
                  &= P(B)P(A \vert B)
            \end{split}
      \end{equation}

      If $A$ and $B$ are independent, then

      \begin{equation}
            P(A \cap B) = P(A)P(B)
      \end{equation}
\end{theorem}

\begin{proof}
      Proof is given from Definition \ref{eq:probability:cond_probability} above.
\end{proof}
\vspace*{0.5cm}

\begin{theorem}[\textbf{The Additive Law of Probability}]
      \label{th:probability:multiple_events:additive}
      The probability of the union of two events $A$ and $B$ is given as

      \begin{equation}
            P(A \cup B) = P(A) + P(B) - P(B \cap A)
      \end{equation}
\end{theorem}

\begin{proof}
      The geometric proof of the additive law of probability is given by the Venn Diagram presented in Figure \ref{fig:probability:multiple_events:additive}. Note that $A \cup B = A \cup ( \bar{A} \cap B)$, where $A$ and $\bar{A} \cap B$ are mutually exclusive events. Furthermore, consider that $B = (\bar{A} \cap B) \cup (A \cap B)$, where $\bar{A} \cap B$ and $A \cap B$ are mutually exclusive. Then $P(A \cup B) = P(A) + P(\bar{A} \cap B)$ and $P(B) = P(\bar{A} \cap B) + P(A \cap B)$. The equality on the right implies that $P(\bar{A} \cap B) = P(B) - P(A \cap B)$. By substituting the expression for $P(\bar{A} \cap B)$ into the expression for $P(A \cup B)$, one can obtain the result $P(A \cup B) = P(A) + P(B) - P(A \cap B)$.
\end{proof}

\begin{figure}[htbp]
      \centering
      \includegraphics[width=0.7\textwidth]{images/additive_law_of_probability_proof.pdf}
      \caption{A Venn-Diagram showing the proof of the additive law of probability for multiple events.}
      \label{fig:probability:multiple_events:additive}
\end{figure}

\index{Bayes Theorem}
\section{Bayes' Theorem}\label{sec:probability:bayes_theorem}

\index{Bayes' theorem}Bayes' theorem, named after Thomas Bayes, describes the probability of an event $A$, based on prior knowledge of conditions that might be related to $A$~\cite{ref:zalta:2015}. In order to derive the formal theorem and proof, first consider Definition \ref{def:probability:bayes_theorem:partition} below~\cite{ref:zalta:2015}.
\\
\begin{definition}
      \label{def:probability:bayes_theorem:partition}
      For some positive integer, $K$, let the sets $B_{1}, B_{2}, \dots, B_{K}$ be such that

      \begin{enumerate}
            \item $S = B_{1} \cup B_{2} \cup \dots \cup B_{K}$
            \item $B_{i} \cap B_{j} = \emptyset$, for $i \neq j$
      \end{enumerate}

      Then the collection of sets $\{B_{1}$, $B_{2}$, $\dots$, $B_{K}\}$ is said to be a partition of S, the union of mutually exclusive subsets.
\end{definition}
\vspace*{0.5cm}

\index{Bayes Theorem}
\begin{theorem}[\textbf{Bayes' theorem}]
      \label{th:probability:bayes_theorem:theorem}
      Assume that $[B_{1}, B_{2}, \dots, B_{K}]$ is a partition of $S$ such that $P(B_{i}) > 0$, for $i = 1,2, \dots, K$ then\\
      \\
      \begin{equation}
            P(B_{j} \vert A) = \frac{P(A \vert B_{j})P(B_{j})}{\sum_{i=1}^{K} P(A \vert B_{i})P(B_{i})}
      \end{equation}
\end{theorem}

\begin{proof}
      The proof follows from the definition of conditional probability as was presented in Section~\ref{sec:probability:cond_probability}.

      \begin{equation}
            \begin{split}
                  P(B_{j} \vert A)
                  &= \frac{P(A \cap B_{j})}{P(A)}\\
                  &= \frac{P(A \vert B_{j})P(B_{j})}{\sum_{i=1}^{K} P(A \vert B_{i})P(B_{i})}
            \end{split}
      \end{equation}
\end{proof}

One of the many applications of \index{Bayes' theorem}Bayes' theorem is to do statistical inference. \index{Bayes' theorem}Bayes' theorem expresses how a degree of belief, expressed as a probability, should rationally change to account for the availability of related evidence.

\section{Probability Distributions}\label{sec:probability:probability_distributions}

Probability distributions are mathematical functions that give the probabilities of the occurrences of different possible outcomes in the experiment. The theory and equations presented in the following sections where all taken from~\citeauthor{ref:wackerly:2014}~\cite{ref:wackerly:2014}.

\subsection{Beta probability distribution}\label{sec:probability:probability_distributions:beta}

The \index{Beta probability distribution}Beta probability distribution is a family of univariate, continuous, probability distributions over some $x$, with support on the interval $[0,1]$~\cite{ref:wackerly:2014}. It is parameterised by two shape parameters $\alpha > 0$, $\alpha \in \mathbb{R}$ and $\beta > 0$, $\beta \in \mathbb{R}$. The \index{Beta probability distribution}Beta probability distribution is denoted as $Beta(\alpha, \beta)$. The \acf{PDF} of the \index{Beta probability distribution}Beta probability distribution is given in Equation~\eqref{eq:probability:probability_distributions:beta:pdf} below.

\begin{equation}
      \label{eq:probability:probability_distributions:beta:pdf}
      P(x \vert \alpha, \beta) = f_{Beta}(x; \alpha, \beta) = \frac{1}{B(\alpha, \beta)} x^{\alpha - 1} (1 - x)^{\beta - 1}
\end{equation}

From Equation~\eqref{eq:probability:probability_distributions:beta:pdf} above, the normalising constant $B(\alpha, \beta)$, is defined below in Equation~\eqref{eq:probability:probability_distributions:beta:norm_const}.

\begin{equation}
      \label{eq:probability:probability_distributions:beta:norm_const}
      B(\alpha, \beta) = \frac{\Gamma(\alpha)\Gamma(\beta)}{\Gamma(\alpha + \beta)}
\end{equation}

In Equation~\eqref{eq:probability:probability_distributions:beta:norm_const}, $\Gamma$ is the Gamma function as defined in Equation~\eqref{eq:probability:probability_distributions:beta:gamma_func} below.

\begin{equation}
      \label{eq:probability:probability_distributions:beta:gamma_func}
      \Gamma(n) = ( n - 1)!
\end{equation}

It should be noted that the Gamma function can also be written as shown in Equation~\eqref{eq:probability:probability_distributions:beta:gamma_func_alt} below.

\begin{equation}
      \label{eq:probability:probability_distributions:beta:gamma_func_alt}
      \Gamma(n+1) = n!
\end{equation}

The control parameters $\alpha$ and $\beta$, determine the shape of the distribution. There exists a special case where $\alpha = \beta$. This is referred to as the \textit{symmetric} \index{Beta probability distribution}Beta probability distribution. In the case where $\alpha = \beta = 1$ the distribution is equivalent to the uniform distribution over all points in its support. The \index{Beta probability distribution}Beta probability distribution for various values of $\alpha$ and $\beta$, including the symmetric version is presented in Figure \ref{fig:probability:probability_distributions:beta}.

\begin{figure}[htbp]
      \begin{subfigure}{0.49\textwidth}
            \centering
            \includegraphics[width=\textwidth]{images/beta_various.pdf}
            \caption{\index{Beta probability distribution}Beta probability distribution}
            \label{fig:probability:probability_distributions:beta_normal}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
            \centering
            \includegraphics[width=\textwidth]{images/beta_cumulative.pdf}
            \caption{Cumulative \index{Beta probability distribution}Beta probability distribution}
            \label{fig:probability:probability_distributions:beta_cumulative}
      \end{subfigure}
      \par\bigskip
      \caption{An illustration of the \index{Beta probability distribution}Beta probability distribution (left)~\cite{ref:beta:2014} as well as the cumulative \index{Beta probability distribution}Beta probability distribution (right)~\cite{ref:cumulativebeta:2014} for various values of $\alpha$ and $\beta$ .}
      \label{fig:probability:probability_distributions:beta}
\end{figure}

The expected value of $x$ is given below in Equation~\eqref{eq:probability:probability_distributions:beta:expected_value}.

\begin{equation}
      \label{eq:probability:probability_distributions:beta:expected_value}
      E[x] = \frac{\alpha}{\alpha + \beta}
\end{equation}

Similarly, the expected value of the natural logarithm of $x$ can be calculated as is shown below in Equation~\eqref{eq:probability:probability_distributions:beta:expected_value_ln}.

\begin{equation}
      \label{eq:probability:probability_distributions:beta:expected_value_ln}
      E[\ln(x)] = \psi({\alpha}) - \psi(\alpha + \beta)
\end{equation}

In Equation~\eqref{eq:probability:probability_distributions:beta:expected_value_ln} above, $\psi$ is the logarithmic derivative of the Gamma function, called the \textit{Digamma} function. The Digamma function is given below in Equation~\eqref{eq:probability:probability_distributions:beta:digamma}.

\begin{equation}
      \label{eq:probability:probability_distributions:beta:digamma}
      \psi(n) = \frac{d}{dn}\ln(\Gamma(n)) = \frac{\Gamma'(n)}{\Gamma(n)}
\end{equation}

Finally, the mode of the distribution is given in Equation~\eqref{eq:probability:probability_distributions:beta:mode} below.

\begin{equation}
      \label{eq:probability:probability_distributions:beta:mode}
      M[x] = E[x] - 1 = \frac{\alpha - 1}{\alpha + \beta - 1}
\end{equation}

\subsection{Dirichlet probability distribution}\label{sec:probability:probability_distributions:dirichlet}

The \index{Dirichlet probability distribution}Dirichlet probability distribution is a family of multivariate continuous probability distributions over some $x$ in $K$ dimensions~\cite{ref:wackerly:2014}. The \index{Dirichlet probability distribution}Dirichlet probability distribution is the multivariate generalization of the \index{Beta probability distribution}Beta probability distribution and is thus sometimes referred to by its alternative name, the multivariate Beta probability distribution. The \index{Dirichlet probability distribution}Dirichlet probability distribution is parameterised by some vector $\vec{\alpha} = (\alpha_{1}$,  $\dots, \alpha_{K})$ $\forall_{k=1}^{K} \alpha_{k} > 0$, $\alpha_{k} \in \mathbb{R}$. Throughout this chapter, the vector notation is dropped for simplicity and readability. $\alpha$ is referred to as the concentration parameter. The \index{Dirichlet probability distribution}Dirichlet probability distribution of order $K \geq 2$ with parameters $\alpha$, denoted $Dir(\alpha)$, has a \acs{PDF} as is given in Equation~\eqref{eq:probability:probability_distributions:dirichlet:pdf} below.
